### 2023-09-14 manuscript.tex
Correct citation [ci skip]
This build is based on
https://github.com/habi/EAWAG-manuscript/commit/1dcd4ef07da5b54f4e19c12f8df6e68aaf709e51.

This commit was created by the following CI build and job:
https://github.com/habi/EAWAG-manuscript/commit/1dcd4ef07da5b54f4e19c12f8df6e68aaf709e51/checks
https://github.com/habi/EAWAG-manuscript/actions/runs/6187112824

\emph{A preprint of this manuscript is available at bioRχiv with the \href{https://doi.org/10.1101/2023.03.30.534917}{doi:10.1101/2023.03.30.534917}.}
\emph{We submitted \href{https://github.com/habi/EAWAG-manuscript/releases/tag/v1.0}{\texttt{v1.0}} (\href{https://habi.github.io/EAWAG-manuscript/v/e1e2ef76a476174a4115937d77457037ddec95df/}{archived here}) of the manuscript to \href{https://journals.plos.org/plosone/}{PLOS ONE}.}
\emph{A revised version \href{https://github.com/habi/EAWAG-manuscript/releases/tag/v1.1}{\texttt{v1.1}} (\href{https://habi.github.io/EAWAG-manuscript/v/c49d0b8ac8f078c226318972b06e8c4a39100bb2/}{archived here}) has been submitted.}
\emph{The version you see here \href{https://github.com/habi/EAWAG-manuscript/compare/df99415...c050bb80f13115ddab1a67428d31c5177ce08ee7\#files_bucket}{is further updated} since re-submission.}

This manuscript
(\href{https://habi.github.io/EAWAG-manuscript/v/c050bb80f13115ddab1a67428d31c5177ce08ee7/}{permalink})
was automatically generated
from \href{https://github.com/habi/EAWAG-manuscript/tree/c050bb80f13115ddab1a67428d31c5177ce08ee7}{habi/EAWAG-manuscript@c050bb8}
on August 22, 2023.

@@

Data on such automatically extracted otoliths, like volume and geometric information like eccentricity and moments of inertia is biologically interesting as the otoliths grow with the age of the fish.
One could help estimate the age of wild fishes using a calibration based on the otolith measurements of a fish of known age.
It is worth noting that the estimation of age in tropical fishes are not as simple as for fishes from temperate regions, where one can distinguish summer and winter layers within the otolith {[}\protect\hyperlink{ref-doi:10.1126ux2fscience.173.4002.112}{\textbf{doi:10.1126/science.173.4002.112?}}{]}.

To the best of our knowledge, no fully automated method for non-destructive otolith extraction from high-resolution tomographic data has been published up to now.
Vasconcelos-Filho et al. {[}\protect\hyperlink{ref-PgMc4mCy}{36}{]} show a potentially objective method to count growth rings in otoliths of six fish species, but performed their study on extracted otoliths ``embedded in a Styrofoam cube'', while our method non-destructively extracts the otoliths from the tomographic data.
Gu et al. {[}\protect\hyperlink{ref-HYXdLQTV}{37}{]} also provide detailed insight into otoliths and combine microtomographic imaging (with unknown voxel size) with scanning electron microscopy.
However, they also mention that the ``otoliths were removed {[}from \emph{Bahaba taipingensis}{]}, cleaned and kept dry for morphologic observation''.

@@

We thank Salome Mwaiko for taking care of the fish collection at Eawag and Mark Charran for helping to find suitable specimens to represent each species.
We are grateful to the \href{https//www.mic.unibe.ch/}{Microscopy Imaging Center} of the University of Bern for the infrastructural support.
We thank the \texttt{manubot} project {[}\protect\hyperlink{ref-YuJbg3zO}{38}{]} for helping us write this manuscript collaboratively.

@@

\leavevmode\vadjust pre{\hypertarget{ref-WVIa4CH0}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{Konings AF, Wisor JM, Stauffer JR Jr. Microcomputed tomography used to link head morphology and observed feeding behavior in cichlids of Lake Malaŵi. Ecol Evol. 2021;11: 4605--4615. doi:\href{https://doi.org/10.1002/ece3.7359}{10.1002/ece3.7359}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-3rxGCEvJ}{}}%
\CSLLeftMargin{10. }%
\CSLRightInline{Summers AP. CT Scans - \#ScanAllFish. 2015 {[}cited 22 Aug 2023{]}. Available: \url{https://osf.io/ecmz4/}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-wsjdcTeC}{}}%
\CSLLeftMargin{15. }%
\CSLRightInline{OpenSCAD. {[}cited 22 Aug 2023{]}. Available: \url{https://openscad.org}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-wIfTDBqd}{}}%
\CSLLeftMargin{23. }%
\CSLRightInline{Image processing with Dask Arrays --- dask-image 2023.08.1+1.g86b56a7.dirty documentation. {[}cited 22 Aug 2023{]}. Available: \url{https://image.dask.org/en/latest/}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-rn6XbmFW}{}}%
\CSLLeftMargin{26. }%
\CSLRightInline{Teem: nrrd: Definition of NRRD File Format. {[}cited 22 Aug 2023{]}. Available: \url{https://teem.sourceforge.net/nrrd/format.html}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-PgMc4mCy}{}}%
\CSLLeftMargin{36. }%
\CSLRightInline{Vasconcelos-Filho JE, Thomsen FSL, Stosic B, Antonino ACD, Duarte DA, Heck RJ, et al. Peeling the Otolith of Fish: Optimal Parameterization for Micro-CT Scanning. Front Mar Sci. 2019;6. doi:\href{https://doi.org/10.3389/fmars.2019.00728}{10.3389/fmars.2019.00728}}

\leavevmode\vadjust pre{\hypertarget{ref-HYXdLQTV}{}}%
\CSLLeftMargin{37. }%
\CSLRightInline{Gu Y-G, Huang H-H, Liang Y, Fang Y, Dai M, Ou Y-J, et al. Micro-CT and SEM investigation of sound absorption structure and chambers in the otoliths of Giant Panda fish species - Chinese Bahaba (Bahaba taipingensis). Micron. 2022;161: 103342. doi:\href{https://doi.org/10.1016/j.micron.2022.103342}{10.1016/j.micron.2022.103342}}

\leavevmode\vadjust pre{\hypertarget{ref-YuJbg3zO}{}}%
\CSLLeftMargin{38. }%
\CSLRightInline{Himmelstein DS, Rubinetti V, Slochower DR, Hu D, Malladi VS, Greene CS, et al. Open collaborative writing with Manubot. Schneidman-Duhovny D, editor. PLoS Comput Biol. 2019;15: e1007128. doi:\href{https://doi.org/10.1371/journal.pcbi.1007128}{10.1371/journal.pcbi.1007128}}

>>>>

\emph{A preprint of this manuscript is available at bioRχiv with the \href{https://doi.org/10.1101/2023.03.30.534917}{doi:10.1101/2023.03.30.534917}.}
\emph{We submitted \href{https://github.com/habi/EAWAG-manuscript/releases/tag/v1.0}{\texttt{v1.0}} (\href{https://habi.github.io/EAWAG-manuscript/v/e1e2ef76a476174a4115937d77457037ddec95df/}{archived here}) of the manuscript to \href{https://journals.plos.org/plosone/}{PLOS ONE}.}
\emph{A revised version \href{https://github.com/habi/EAWAG-manuscript/releases/tag/v1.1}{\texttt{v1.1}} (\href{https://habi.github.io/EAWAG-manuscript/v/c49d0b8ac8f078c226318972b06e8c4a39100bb2/}{archived here}) has been submitted.}
\emph{The version you see here \href{https://github.com/habi/EAWAG-manuscript/compare/df99415...1dcd4ef07da5b54f4e19c12f8df6e68aaf709e51\#files_bucket}{is further updated} since re-submission.}

This manuscript
(\href{https://habi.github.io/EAWAG-manuscript/v/1dcd4ef07da5b54f4e19c12f8df6e68aaf709e51/}{permalink})
was automatically generated
from \href{https://github.com/habi/EAWAG-manuscript/tree/1dcd4ef07da5b54f4e19c12f8df6e68aaf709e51}{habi/EAWAG-manuscript@1dcd4ef}
on September 14, 2023.

@@

Data on such automatically extracted otoliths, like volume and geometric information like eccentricity and moments of inertia is biologically interesting as the otoliths grow with the age of the fish.
One could help estimate the age of wild fishes using a calibration based on the otolith measurements of a fish of known age.
It is worth noting that the estimation of age in tropical fishes are not as simple as for fishes from temperate regions, where one can distinguish summer and winter layers within the otolith {[}\protect\hyperlink{ref-Iy1X7kEU}{36}{]}.

To the best of our knowledge, no fully automated method for non-destructive otolith extraction from high-resolution tomographic data has been published up to now.
Vasconcelos-Filho et al. {[}\protect\hyperlink{ref-PgMc4mCy}{37}{]} show a potentially objective method to count growth rings in otoliths of six fish species, but performed their study on extracted otoliths ``embedded in a Styrofoam cube'', while our method non-destructively extracts the otoliths from the tomographic data.
Gu et al. {[}\protect\hyperlink{ref-HYXdLQTV}{38}{]} also provide detailed insight into otoliths and combine microtomographic imaging (with unknown voxel size) with scanning electron microscopy.
However, they also mention that the ``otoliths were removed {[}from \emph{Bahaba taipingensis}{]}, cleaned and kept dry for morphologic observation''.

@@

We thank Salome Mwaiko for taking care of the fish collection at Eawag and Mark Charran for helping to find suitable specimens to represent each species.
We are grateful to the \href{https//www.mic.unibe.ch/}{Microscopy Imaging Center} of the University of Bern for the infrastructural support.
We thank the \texttt{manubot} project {[}\protect\hyperlink{ref-YuJbg3zO}{39}{]} for helping us write this manuscript collaboratively.

@@

\leavevmode\vadjust pre{\hypertarget{ref-WVIa4CH0}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{Konings AF, Wisor JM, Stauffer JR Jr. Microcomputed tomography used to link head morphology and observed feeding behavior in cichlids of Lake Malaŵi. Ecology and Evolution. 2021;11: 4605--4615. doi:\href{https://doi.org/10.1002/ece3.7359}{10.1002/ece3.7359}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-3rxGCEvJ}{}}%
\CSLLeftMargin{10. }%
\CSLRightInline{Available: \url{https://osf.io/ecmz4}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-wsjdcTeC}{}}%
\CSLLeftMargin{15. }%
\CSLRightInline{OpenSCAD. {[}cited 14 Sep 2023{]}. Available: \url{https://openscad.org}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-wIfTDBqd}{}}%
\CSLLeftMargin{23. }%
\CSLRightInline{Image processing with Dask Arrays --- dask-image 2023.08.1+3.gc52eec3.dirty documentation. {[}cited 14 Sep 2023{]}. Available: \url{https://image.dask.org/en/latest/}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-rn6XbmFW}{}}%
\CSLLeftMargin{26. }%
\CSLRightInline{Teem: nrrd: Definition of NRRD File Format. {[}cited 14 Sep 2023{]}. Available: \url{https://teem.sourceforge.net/nrrd/format.html}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Iy1X7kEU}{}}%
\CSLLeftMargin{36. }%
\CSLRightInline{Pannella G. Fish Otoliths: Daily Growth Layers and Periodical Patterns. Science. 1971;173: 1124--1127. doi:\href{https://doi.org/10.1126/science.173.4002.1124}{10.1126/science.173.4002.1124}}

\leavevmode\vadjust pre{\hypertarget{ref-PgMc4mCy}{}}%
\CSLLeftMargin{37. }%
\CSLRightInline{Vasconcelos-Filho JE, Thomsen FSL, Stosic B, Antonino ACD, Duarte DA, Heck RJ, et al. Peeling the Otolith of Fish: Optimal Parameterization for Micro-CT Scanning. Front Mar Sci. 2019;6. doi:\href{https://doi.org/10.3389/fmars.2019.00728}{10.3389/fmars.2019.00728}}

\leavevmode\vadjust pre{\hypertarget{ref-HYXdLQTV}{}}%
\CSLLeftMargin{38. }%
\CSLRightInline{Gu Y-G, Huang H-H, Liang Y, Fang Y, Dai M, Ou Y-J, et al. Micro-CT and SEM investigation of sound absorption structure and chambers in the otoliths of Giant Panda fish species - Chinese Bahaba (Bahaba taipingensis). Micron. 2022;161: 103342. doi:\href{https://doi.org/10.1016/j.micron.2022.103342}{10.1016/j.micron.2022.103342}}

\leavevmode\vadjust pre{\hypertarget{ref-YuJbg3zO}{}}%
\CSLLeftMargin{39. }%
\CSLRightInline{Himmelstein DS, Rubinetti V, Slochower DR, Hu D, Malladi VS, Greene CS, et al. Open collaborative writing with Manubot. Schneidman-Duhovny D, editor. PLoS Comput Biol. 2019;15: e1007128. doi:\href{https://doi.org/10.1371/journal.pcbi.1007128}{10.1371/journal.pcbi.1007128}}
